---
title: "BIOL432_Assignment1"
author: "SiruiZHAO"
date: "2023-01-11"
output: html_document
---
[my GitHub repository](https://github.com/siruiZHAO/BIOL432_Assignment1)

First, load library
```{r message=FALSE}
library(ggplot2)
library(dplyr)
```
Then, we need to source two R scripts. 

1). dataGenerato.R generate 4 varuables, organism, limb width, limb length and observer, and combine them into a data frame. 

2). volumeEstimato.R add one variable called limb volume into the data frame. The limb volume calculated from the v = pi * (limb_width/2)^2 * limb_length
```{r}
source("dataGenerato.R", local = knitr::knit_global())
source("volumeEstimato.R", local = knitr::knit_global())
```

1. Load the revised measurements.csv and library
```{r}
mydata <- read.csv("measurements.csv")
```

2. Use dplyr to sort the data by species, then by observer, then by limb volume
```{r}
mydata <- mydata %>% #use pipe in dplyr 
  arrange(organism, Observer, limb_volume) # use arrange to sort column in order
head(mydata)
```

3. Create a table showing the average volume for each species
```{r}
avg_volume <- mydata %>%     #select data frame
  group_by(organism) %>%   #group rows by column values in the DataFrame
  summarise(mean_volum = mean(limb_volume)) #calculate mean of limb volume
avg_volume
```

4. Create a table showing the number of observations (i.e. rows) for each combination of species and observer.
```{r}
observation <- mydata %>%     #select data frame
  group_by(organism, Observer) %>%  #group rows by two column values 
  summarise(observations = n(), .group = "drop") %>%  #count frequency
  as.data.frame()
observation
```

5. Create a single box plot (or similar) to compare key features of the distributions of volumes for each species.
```{r}
#use ggplot2 to create box plot
plot1 <- ggplot(mydata, aes(y = limb_volume, fill = organism)) + geom_boxplot() + labs(x = "Organism Scientific Name", y = "Limb Volume(cm^3)")
plot1
```

6. Create a multi-panel plot showing frequency histograms showing the distribution of limb volume for each species. Note that this may require you to add code to you R markdown file to change the data from the wide to long format.
```{r}
#use facet_wrap() in ggplot2 to create multi-panel plot
ggplot(mydata, aes(x = limb_volume)) +
  geom_histogram() +
  facet_wrap(vars(organism)) + labs(x = "Limb Volume (cm^3)", y = "Frequency")
```
Thanks!
